A formal approach but practical approach in train regulation

### 2025-06-12 paper/main.tex
regulators rules

Now, we come to the train model. Each train has an identifier. In other words, we have a set of trains $\trains = \{1, \ldots, k\}$. Each train has a position on the network. The position is the index of the block it stays in or the last block it was in. Next, a train has a direction in $\directions$. The direction does not mean the train is running---it can be blocked by some red signal---but that its next move will be in that direction. Finally, a train has duties. Duties are given as a sequence of orders following the grammar:
$$  seq ::= \su(dir,pos) $$
with $dir \in \directions$ and $pos \in \blocks$. A duty $\su(dir, pos)$ states that the train direction is set to $dir$ and that the train runs until it reaches $pos$.  The sequence may be empty.

@@

\subsection{Modélisation d'une liste d'odre}
La modélisation de notre liste d'ordre doit prendre en compte les besoins et contraintes évoqués plus haut.
Pour cela, nous avons choisi de la représenter comme cela : $O = \{\Gamma, R\}$, avec $\Gamma$ l'ensemble des trains et $R$ notre régulateur.
Chaque train peut être représenté par un identifiant unique, une position, une direction et son programme : $t = (id, pos, dir, P)$, avec $t \in \Gamma$.
Le régulateur, portant une grande partie de la logique du modèle, se représente de manière plus complexe : $R = (E,T,S,W,G,H,TL,B)$.
%##########
%Définit ce qu'est un event à ce moment ou plus haut ?
%##########
Nous ne détaillerons pas l'ensemble des éléments de $R$ ici, mais nous pouvons en donner un aperçu rapide : $E$ contient l'ensemble des events,
$T$ le token courant pour chaque ressource et $H$ la dernière position connue de chaque train par le régulateur.

>>>>

Now, we come to the train model. Each train has an identifier. In other words, we have a set of trains $\trains = \{1, \ldots, k\}$. Each train has a position on the network. The position is the index of the block it stays in or the last block it was in. Next, a train has a direction in $\directions$. The direction does not mean the train is running---it can be blocked by some red signal---but that its next move will be in that direction. Finally, a train has orders. Orders are given as a sequence of orders following the grammar:
$$  seq ::= \su(dir,pos) $$
with $dir \in \directions$ and $pos \in \blocks$. An order $\su(dir, pos)$ states that the train direction is set to $dir$ and that the train runs until it reaches $pos$.  The sequence may be empty.

To sum up, trains can be viewed as 4-tuples: $t = (id, pos, dir, P)$ with $id$ the train's identifier, $pos$ its position, $dir$ its direction and $P$ the sequence of orders. The description of all trains is then a sequence $\Gamma = (t_1, \ldots, t_k)$ with $t_1$ the 4-tuple corresponding to train 1, up to $t_k$ that is the 4-tuple corresponding to train $k$.

The regulator is characterized as follows. First, it has a general program, next mentioned as $E$. The program contains the plan of the regulator. The regulator will follow its rule. A program is given as a sequence of statements. Those follow the grammar:

\begin{grammar}[Regulator's rules][h][regulateur]
    \firstcasesubtil{stmt}{ev(id, n) : atoms}{}
    \firstcasesubtil{atoms}{incr(j)}{}
    \otherform{att(j, value_{j})}{}
    \otherform{turn(id_{switch}, state)}{}
    \otherform{auth}{}
    \otherform{atoms \gralt \varepsilon}{}
\end{grammar}

@@

\subsection{Modélisation d'une liste d'odre}
%La modélisation de notre liste d'ordre doit prendre en compte les besoins et contraintes évoqués plus haut.
%Pour cela, nous avons choisi de la représenter comme cela : $O = \{\Gamma, R\}$, avec $\Gamma$ l'ensemble des trains et $R$ notre régulateur.
%Chaque train peut être représenté par un identifiant unique, une position, une direction et son programme : $t = (id, pos, dir, P)$, avec $t \in \Gamma$.
Le régulateur, portant une grande partie de la logique du modèle, se représente de manière plus complexe : $R = (E,T,S,W,G,H,TL,B)$.
%##########
%Définit ce qu'est un event à ce moment ou plus haut ?
%##########
Nous ne détaillerons pas l'ensemble des éléments de $R$ ici, mais nous pouvons en donner un aperçu rapide : $E$ contient l'ensemble des events,
$T$ le token courant pour chaque ressource et $H$ la dernière position connue de chaque train par le régulateur.
